### 2016-11-10 paper/modes.tex
Updated RQs
Updated parallel modes

\section{Parallel Execution of Test Suites}

@@

This paper focuses at lower-level parallelism, obtained with parallel
computation at cpus and threads in a given machine.  This kind of
parallelism is enabled through build systems and testing frameworks.
Given the proliferation of multicore machines it is not surprising
that testing frameworks provide today support for parallel test
execution with the goal of running tests more
efficiently~\cite{junit-org,testng,nunit}.  Build systems wrap
parallel features offered by testing frameworks to facilitate the use
of such technology\cite{maven-surefire-plugin}.  In the following we
elaborate relevant features of these components focusing on Java,
Maven, and JUnit but the discussion can be generalized to other
languages, build systems, and testing frameworks.

Note that lower-level parallelism is a complement to higher-level
parallelism that could be enabled through server farms.  For the
scenario of large IT organizations, lower-level parallelization
schemes could leverage the computing power of server nodes in addition
to the aggregate processing power of the farm.  Lower-level
parallelism fits particularly well smaller organizations (/projects)
with relatively high testing costs but lower budgets.
\Comment{Unfortunately, these solutions can't be used out-of-the-box:
  unrestricted parallel execution of tests can produce
  non-deterministic results as developers typically do not provision
  protection to concurrent accesses originated from arbitrary program
  points (\ie{}, tests).  We refer to this problem as Parallel
  Execution Flakiness (\pef{}).}

@@

\Comment{
To illustrate the importance of parallel test execution and
the problem of \pef{} let's consider the case of the ``core'' module
from the Apache Camel project~\cite{apache-camel-web}.  This module
contains 5,679 test cases, declared in 2,356 test classes.  We ran
those tests in a machine with 16GB of memory and 8 virtual CPUs (4
cores with 2 native threads each).  Sequential test execution takes
24m50s to run this test set.  Execution of the same test set takes
2m28s when we configured parallel execution to fork a JVM per CPU and
execute test classes, uniquely allocated to that JVM, sequentially but
running test methods from each class in separate threads.  Note that
this is an order of magnitude speedup (10.07x)\Fix{Need to understand
why this is 10x as opposed to something closer to 7x - I didn't get
your concern here}.  Unfortunately, due to \pef{}, $\sim$2\% (114 of
5,679) of the tests fail when executed in parallel.  It is important
to notice that the ratio of failures varies with the project as it
depends on factors such as length of test cases and amount of shared
state across tests.

\pef{} is an important obstacle to enable parallel test execution.
Conceptually, higher parallelization can result in higher chances of
concurrency-related problems.  It is important to execute tests
efficiently without sacrificing reliability
\Fix{gap between pars?? What gap?}

}

@@

The list below shortly describes the choices to control parallelism
within each JVM.

@@

\begin{itemize}
\item \textbf{Sequential (\Seq).}~This configuration results in
  sequential execution of test classes and their methods, according to
  the order defined in the testing framework\Fix{cite}.  For JUnit
  this is the default configuration.
\item \textbf{Sequential classes; parallel methods
  (\ParClassSeqMeth{}).} This configuration results in the sequential
  execution of the test classes assigned to the JVM.  However, within
  each class, test methods run in parallel.\Fix{are there options to
    limit number of threads or a thread per test method?  Do they use
    thread pools?  etc....}
\item \textbf{Parallel classes; sequential methods
  (\SeqClassParMeth{}).}  This configuration results in the parallel
  execution of test classes but test methods of any given class are
  executed sequentially.\Fix{one thread per class?  if not, what
    classes are scheduled to execute first?}
\item \textbf{Parallel methods (\ParClassParMeth).} This configuration
  results in parallel execution of any given test method of any class.
\end{itemize}

>>>>

\section{Parallel Execution of Test Suites}
\label{sec:modes}

@@

This paper focuses at lower-level parallelism, obtained with parallel
computation at CPUs and threads in a given machine.  This kind of
parallelism is enabled through build systems and testing frameworks.
Given the proliferation of multicore machines it is not surprising
that testing frameworks provide today support for parallel test
execution with the goal of running tests more
efficiently~\cite{junit-org,testng,nunit}.  Build systems wrap
parallel features offered by testing frameworks to facilitate the use
of such technology\cite{maven-surefire-plugin}.  In the following we
elaborate relevant features of these components focusing on Java,
Maven, and JUnit but the discussion can be generalized to other
languages, build systems, and testing frameworks.

Note that lower-level parallelism is a complement to higher-level
parallelism that could be enabled through server farms.  For the
scenario of large IT organizations, lower-level parallelization
schemes could leverage the computing power of server nodes in addition
to the aggregate processing power of the farm.  Lower-level
parallelism fits particularly well smaller organizations (/projects)
with relatively high testing costs but lower budgets.
\Comment{
    Unfortunately, these solutions can't be used out-of-the-box:
    unrestricted parallel execution of tests can produce
    non-deterministic results as developers typically do not provision
    protection to concurrent accesses originated from arbitrary program
    points (\ie{}, tests).  We refer to this problem as Parallel
    Execution Flakiness (\pef{}).

@@

    To illustrate the importance of parallel test execution and
    the problem of \pef{} let's consider the case of the ``core'' module
    from the Apache Camel project~\cite{apache-camel-web}.  This module
    contains 5,679 test cases, declared in 2,356 test classes.  We ran
    those tests in a machine with 16GB of memory and 8 virtual CPUs (4
    cores with 2 native threads each).  Sequential test execution takes
    24m50s to run this test set.  Execution of the same test set takes
    2m28s when we configured parallel execution to fork a JVM per CPU and
    execute test classes, uniquely allocated to that JVM, sequentially but
    running test methods from each class in separate threads.  Note that
    this is an order of magnitude speedup (10.07x)\Fix{Need to understand
    why this is 10x as opposed to something closer to 7x - I didn't get
    your concern here}.  Unfortunately, due to \pef{}, $\sim$2\% (114 of
    5,679) of the tests fail when executed in parallel.  It is important
    to notice that the ratio of failures varies with the project as it
    depends on factors such as length of test cases and amount of shared
    state across tests.

    \pef{} is an important obstacle to enable parallel test execution.
    Conceptually, higher parallelization can result in higher chances of
    concurrency-related problems.  It is important to execute tests
    efficiently without sacrificing reliability
    \Fix{gap between pars?? What gap?}
}

@@

The list below shortly describes the choices to control parallelism
within a JVM.

@@

\begin{itemize}
\item \textbf{Sequential (\Seq).}~This configuration results in
    sequential execution of test classes and their methods, according
    to a given order defined in the testing framework (\eg, the order
    of methods returned by Java Reflection
    API)~\cite{junit-test-order}. For JUnit, this is the default
    configuration.

\item \textbf{Sequential classes; parallel methods
  (\ParClassSeqMeth{}).} This configuration results in the sequential
  execution of the test classes assigned to the JVM.  However, within
  each class, test methods run in parallel. In JUnit,
  \emph{ParallelComputer} provides support to parallel execution; it
  instantiates a test runner with a \emph{ExecutorService} from the
  Java Concurrent API. Each test method is executed in a separated
  thread as the thread pool creates new threads on-demand and reuse
  them if any thread is available.

\Comment{http://junit.org/junit4/javadoc/latest/src-html/org/junit/experimental/ParallelComputer.html#line.14}

\item \textbf{Parallel classes; sequential methods
  (\SeqClassParMeth{}).}  This configuration results in the parallel
  execution of test classes but test methods of any given class are
  executed sequentially. As in \ParClassSeqMeth, the runner has a
  cached thread poll. However, instead of executing test methods from
  a single test class, the test suite runner executes classes in
  separated threads and methods in sequence within each thread.

\item \textbf{Parallel methods (\ParClassParMeth).} This configuration
  results in parallel execution of any given test method of any class.
  \Jbc{I need to confirm this but AFAIK this is how it happens}
  It is the union of \ParClassSeqMeth and \SeqClassParMeth: there is a
  thread pool for execution classes on separated threads and within
  each test class, there is a runner with a \emph{ExecutorService} to
  run test methods in parallel.
\end{itemize}

\Jbc{I have to understand better how Surefire executes tests. I'm
    afraid it has its own wrapper classes and reuse JUnit core
    functionalities. I'm saying this because JUnit does not offer a
    finer control over thread numbers while this configuration is
possible with Surefire} For example, Figure~\ref{fig:surefire}
illustrates how to configure Maven to execute classes sequentially
(default behavior) and test methods in parallel. Although the setting
are configured on the build file (\eg, \emph{pom.xml} on Maven}, these
settings are forwarded to the underlying testing framework (\eg,
JUnit). In case of JUnit,

\begin{figure}[h!]
    \centering
    \lstset{language=XML}
    \begin{lstlisting}
    <plugin>
        <groupId>org.apache.maven.plugins</groupId>
        <artifactId>maven-surefire-plugin</artifactId>
        <configuration>
            <parallel>methods</parallel>
            <threadCount>2</threadCount>
            <perCoreThreadCount>true</perCoreThreadCount>
        </configuration>
    </plugin>
    \end{lstlisting}
    \caption{\label{fig:surefire}\Fix{...}}
\end{figure}
